DATA 621 - Business Analytics and Data Mining

### 2020-09-24 HW2/HW2-Group2.rmd
Finished through 7

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, error=FALSE, warning=FALSE, message=FALSE)
```

# 1. The Data
```{r}
library(tidyverse)
# urlRemote  = "https://raw.githubusercontent.com/"
# pathGithub = "aadler/DT621_Fall2020_Group2/master/HW1/data/"
# fileTrain   = "classification-output-data.csv"
# data = paste0(urlRemote, pathGithub, fileTrain) %>% read.csv()

@@

```{r}
cat("----------------------------------------------------------------------------
                 |           Target          |
Confusion Matrix |---------------------------|
                 |   Positive  |   Negative  |
-----------------|-------------|-------------|------------------------------
      | Positive |a (True Pos) |b (False Pos)| Positive Predictive | a/(a+b)
Model |----------|-------------|-------------|------------------------------
      | Negative |c (False Neg)|d (True Neg) | Negative Predictive | d/(c+d)
-----------------|-------------|-------------|------------------------------
                 | Sensitivity | Specificity |                    
                 |-------------|-------------|  Accuracy = (a+d)/(a+b+c+d)
                 |   a/(a+c)   |   d/(b+d)   | 
----------------------------------------------------------------------------")
```

@@

```{r data_table}
class_data = data %>% select(class, pred_class = scored.class,
                             pred_prob = scored.probability) 
table(class_data[,1:2]) %>% knitr::kable(caption = "Raw Confusion Matrix")
```

@@

```{r accuracy, echo = TRUE}
ACC <- function(df) {
  if (!is.data.frame(df)) {
    stop("Data set is not a dataframe. Try as.data.frame(dataset).")
  } else {
    dt <- table(df[, 1:2])
    sum(diag(dt)) / sum(dt)
  }
}
```

@@

```{r error, echo = TRUE}
CER <- function(df) {
  if (!is.data.frame(df)) {
    stop("Data set is not a dataframe. Try as.data.frame(dataset).")
  } else {
    dt <- table(df[, 1:2])
    (sum(dt) - sum(diag(dt))) / sum(dt)
  }
}
```

@@

```{r precision, echo = TRUE}
precision = function(df) {
  if (is.data.frame(df) == FALSE) {
    return("Data set is not a dataframe. Try as.data.frame(dataset).")
  } else {
    dt = as.data.frame(table(df[, 1:2]))
    FP = dt %>% filter(class == 0 & pred_class == 1) %>% select(Freq)
    TP = dt %>% filter(class == 1 & pred_class == 1) %>% select(Freq)
    return(TP / (TP + FP))
  }
}
```

```{r precision_check}
sprintf("The precision of the predictions for the classification data set is %0.3f.", 
precision(class_data))
```

### Sensitivity of the Predictions

Function `sensitivity` that takes the data set as a dataframe, with actual and predicted classifications identified, and returns the sensitivity of the predictions. Sensitivity is also known as recall. The formula is:

\[sensitivity = \frac{TP}{TP + FN}\]

```{r sensitivity, echo = TRUE}
sensitivity = function(df) {
  if (is.data.frame(df) == FALSE) {
    return("Data set is not a dataframe. Try as.data.frame(dataset).")
  } else {
    dt = as.data.frame(table(df[, 1:2]))
    FN = dt %>% filter(class == 1 & pred_class == 0) %>% select(Freq)
    TP = dt %>% filter(class == 1 & pred_class == 1) %>% select(Freq)
    return(TP / (TP + FN))
  }
}
```

```{r sensitivity_check}
sprintf("The sensitivity of the predictions for the classification data set is %0.3f.", 
sensitivity(class_data))
```

### Specificity of the Predictions

Function `specificity` that takes the data set as a dataframe, with actual and predicted classifications identified, and returns the specificity of the predictions. The formula is:

\[specificity = \frac{TN}{TN + FP}\]

```{r specificity, echo = TRUE}
specificity = function(df) {
  if (is.data.frame(df) == FALSE) {
    return("Data set is not a dataframe. Try as.data.frame(dataset).")
  } else {
    dt = as.data.frame(table(df[, 1:2]))
    TN = dt %>% filter(class == 0 & pred_class == 0) %>% select(Freq)
    FP = dt %>% filter(class == 0 & pred_class == 1) %>% select(Freq)
    return(TN / (TN + FP))
  }
}
```

```{r specificity_check}
sprintf("The specificity of the predictions for the classification data set is %0.3f.", 
specificity(class_data))
```

@@

```{r f1_score, echo = TRUE}
f1_score = function(df) {
  if (is.data.frame(df) == FALSE) {
    return("Data set is not a dataframe. Try as.data.frame(dataset).")
  } else {
    return((2*precision(df)*sensitivity(df)) / (precision(df) + sensitivity(df)))
  }
}
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, error=FALSE, warning=FALSE, message=FALSE)
```

# 0. Preamble
***For all the following answers, it is expected that we are only dealing with***
***binary classification in which negative is coded as 0 and positive is coded***
***as 1. It is not difficult to expand the functions to respond to more general***
***cases, but it is an unnecessary complication for this homework assignment.***

# 1. The Data
```{r loadData, echo=FALSE}
library(tidyverse)
# urlRemote  = "https://raw.githubusercontent.com/"
# pathGithub = "aadler/DT621_Fall2020_Group2/master/HW1/data/"
# fileTrain   = "classification-output-data.csv"
# data = paste0(urlRemote, pathGithub, fileTrain) %>% read.csv()

@@

```{r confMatrixDesc, echo=FALSE}
cat("----------------------------------------------------------------------------
                 |           Target          |
Confusion Matrix |---------------------------|
                 |   Positive  |   Negative  |
-----------------|-------------|-------------|------------------------------
      | Positive |a (True Pos) |b (False Pos)| Positive Predictive | a/(a+b)
Model |----------|-------------|-------------|------------------------------
      | Negative |c (False Neg)|d (True Neg) | Negative Predictive | d/(c+d)
-----------------|-------------|-------------|------------------------------
                 | Sensitivity | Specificity |                    
                 |-------------|-------------|  Accuracy = (a+d)/(a+b+c+d)
                 |   a/(a+c)   |   d/(b+d)   | 
----------------------------------------------------------------------------")
```

@@

```{r dataCM, echo=FALSE}
class_data = data %>% select(class, pred_class = scored.class,
                             pred_prob = scored.probability) 
table(class_data[,1:2]) %>% knitr::kable(caption = "Raw Confusion Matrix")
```

@@

```{r accuracy}
ACC <- function(df) {
  if (!is.data.frame(df)) {
    stop("Data set is not a dataframe. Try as.data.frame(dataset).")
  } else {
    dt <- table(df[, 1:2])
    sum(diag(dt)) / sum(dt)
  }
}
```

@@

```{r error}
CER <- function(df) {
  if (!is.data.frame(df)) {
    stop("Data set is not a dataframe. Try as.data.frame(dataset).")
  } else {
    dt <- table(df[, 1:2])
    (sum(dt) - sum(diag(dt))) / sum(dt)
  }
}
```

@@

```{r precision}
PRC <- function(df) {
  if (!is.data.frame(df)) {
    stop("Data set is not a dataframe. Try as.data.frame(dataset).")
  } else {
    dt <- table(df[, 1:2])
    dt[2, 2] / sum(dt[, 2])
  }
}
```

For this dataset the precision is
`r prettyNum(PRC(class_data) * 100, digits = 5L)`%.

# 6. Sensitivity or Recall
The function `SNS` below takes as input a data frame whose first two columns
are the true class and the modeled class respectively. It returns the
sensitivity of the predictions using the following formula:

\[
\textrm{Sensitivity} = \frac{TP}{TP + FN}
\]

```{r sensitivity}
SNS <- function(df) {
  if (!is.data.frame(df)) {
    stop("Data set is not a dataframe. Try as.data.frame(dataset).")
  } else {
    dt <- table(df[, 1:2])
    dt[2, 2] / sum(dt[2, ])
  }
}
```

For this dataset the sensitivity is
`r prettyNum(SNS(class_data) * 100, digits = 5L)`%.

# 7. Specificity
The function `SPC` below takes as input a data frame whose first two columns
are the true class and the modeled class respectively. It returns the
specificity of the predictions using the following formula:

\[
\textrm{Specificity} = \frac{TN}{TN + FP}
\]

```{r specificity}
SPC <- function(df) {
  if (!is.data.frame(df)) {
    stop("Data set is not a dataframe. Try as.data.frame(dataset).")
  } else {
    dt <- table(df[, 1:2])
    dt[1, 1] / sum(dt[1, ])
  }
}
```

For this dataset the specificity is
`r prettyNum(SPC(class_data) * 100, digits = 5L)`%.

@@

```{r f1_score, echo = TRUE}
f1_score = function(df) {
  if (is.data.frame(df) == FALSE) {
    return("Data set is not a dataframe. Try as.data.frame(dataset).")
  } else {
    return((2 * PRC(df) * SNS(df)) / (PRC(df) + SNS(df)))
  }
}
```
